Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance Settings

### 2020-01-24 Manuscript/Figures/figure_legends.Rmd
edits

**Figure 4: Assessing top ranked transcription factors in TAD boundary prediction.** (A) Barplots comparing performances of models using specific classes of functional genomic elements as the feature space including histone modifications (HM; red), cis-regulated elements (CRE; green), transcription factor binding sites (TFBS; blue), in addition to a model with a feature space containing all three classes (Combined; purple). (B) Recursive feature elimination analysis used to reduce the number of TFBS to include in the optimally build random forest classifier. Error bars represent standard deviation from the mean MCC across each chromosome. (C) Clustered heatmap of predictive importances for the union of the top 8 most predictive TFBS across chromosomes.

>>>>

**Figure 4. Assessing top ranked transcription factors in TAD boundary prediction for GM12878.** (A) Barplots comparing performances of models using specific classes of functional genomic elements as the feature space including histone modifications (HM; red), cis-regulated elements (CRE; green), transcription factor binding sites (TFBS; blue), in addition to a model with a feature space containing all three classes (Combined; purple). (B) Recursive feature elimination analysis used to reduce the number of TFBS to include in the optimally build random forest classifier. Error bars represent standard deviation from the mean MCC across each chromosome. (C) Clustered heatmap of predictive importances for the union of the top 8 most predictive TFBS across chromosomes.

**Figure 5. Predicted TAD boundaries are enriched with known molecular drivers of the 3D chromatin architecture.** (A) Visualizing locations of predicted and called TAD boundaries (chr 22: 17390000-18640000; GM12878). The upper triangular contact map shows the normalized number of contacts within the specified window using the JUICER toolbox. Called TADs are outlined in blue. The blue vertical lines represent a called TAD boundary identified by the ARROWHEAD algorithm. The black line represents the predicted probability of each base pair being a TAD boundary. Regions of adjacent base pairs with probability of 1 are represented by the green regions along the x-axis. The green vertical lines represent the subsequent base pair predicted to be a TAD boundary using medoid partitioning. <!--(B) Enrichment heatmaps illustrate the average signal strength of ChIP-seq peak regions defined by SMC3, ZNF143, CTCF, and RAD21 around the centers of 10 kb flanked predicted and called TAD boundaries.-->

### 2020-01-24 Manuscript/Supplementary/supplementary_legends.Rmd
edits

>>>>

**Supplementary Figure 4. Assessing top ranked transcription factors in TAD boundary prediction for K562.** (A) Barplots comparing performances of models using specific classes of functional genomic elements as the feature space including histone modifications (HM; red), cis-regulated elements (CRE; green), transcription factor binding sites (TFBS; blue), in addition to a model with a feature space containing all three classes (Combined; purple). (B) Recursive feature elimination analysis used to reduce the number of TFBS to include in the optimally build random forest classifier. Error bars represent standard deviation from the mean MCC across each chromosome. (C) Clustered heatmap of predictive importances for the union of the top 8 most predictive TFBS across chromosomes.

### 2020-01-24 Manuscript/manuscript.Rmd
edits

Biological replicate Hi-C data were downloaded from GEO GSE63525 [@rao20143d] (Supplementary Table 1) for the GM12878 and K562 cell lines. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from Hi-C data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the ARROWHEAD algorithm [@durand2016juicebox]. TAD boundaries called on chromosome 9 were omitted from all downstream analyses, for all resolutions and cell lines, due to the sparsity of the contact matrix at the 10 kb resolution for the K562 cell line.

Genomic annotation data in the form of cell type-specific cis-regulatory elements (CRE), histone modifications (HM), and transcription factor binding sites (TFBS) were obtained from the UCSC Genome Browser Database (GRCh37/hg19 human genome assembly) [@Tyner:2017aa] (Supplementary Table 2).

## Machine learning framework for TAD boundary prediction

@@

where TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary (false negatives). Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. [???include reason for using mcc, why not others?] <!--Additional performance metrics including AUROC, accuracy, sensitivity, and specificity were reported in the supplementary materials.-->

@@

Many genomic annotations, notably architectural proteins, tend to exhibit an extensive pattern of colocalization (correlation) [@van2014insulator]. To suitably reduce the feature space and prevent overfitting, we implemented recursive feature elimination to estimate the optimal number of necessary features to include in the model. The best performing subset was identified, ranging from 2 to the maximum number of features by powers of 2. We then aggregated the predictive importance of the union of the optimal set of features across chromosomes using the mean decrease in node impurity to determine the most common and top ranked annotations for predicting TAD boundaries. [??? TBD, at which stage it was done]

@@

Next, we aimed to investigate whether we could alleviate resolution limitations that plague conventional TAD calling algorithms. To do so, we extended our optimally built TAD prediction models to obtain predictions at base pair (bp) resolution. Naturally, when dealing with such fine resolution, there will exist collections of spatially close base pairs with similar predictor profiles (i.e. similar distances), and will thus yield similar predictions. To guard against inflating the number of predicted TAD boundaries we developed a technique that combines both hierarchical clustering and partitioning techniques to identify more precise TAD boundaries throughout the linear genome.

@@

We developed a computational framework, refered to as **TADRF**, aimed at determining the optimal set of data level characteristics necessary for TAD boundary prediction using Random Forest classification models. Our framework is resolution-specific and uses called TAD boundaries as ground truth to establish TAD-boundary and non-TAD boundary regions. Under our framework, we binned the linear genome at intervals specific to the Hi-C data resolution used to call TADs. TAD boundary regions were defined as regions on the linear genome where a genomic bin contained a called TAD boundary ($Y = 1$), while non-TAD boundary regions were defined as genomic bins that did not contain a TAD boundary ($Y = 0$) (Figure 1 A). Under this construction, the number of non-TAD boundary regions are exceed the number of TAD boundary regions. To cope with the imbalance among the class of genomic bins, we evaluated three re-sampling techniques (in addition to no re-sampling) before predicting TAD boundaries. **Random over-sampling** (ROS) was defined as sampling with replacement from TAD boundary regions. **Random under-sampling** (RUS) was defined as sampling with replacement from non-TAD boundary regions. Lastly, we considered **Synthetic minority over-sampling technique** (SMOTE), which is a combination of both random over- and under-sampling [@chawla2002smote].

Various one-dimensional cell line-specific ChIP-seq tracks (genomic elements) were used to construct the feature space for the random forests including cis-regulatory elements, histone modifications, and transcription factor binding sites. Four different feature engineering procedures were developed for each genomic element (Figure 1 B). **Peak Signal** was defined as the [???include definition]. Counts of overlaps (**Overlap Counts**, OC) were defined as the total number of genomic annotation regions that overlapped within a given genomic bin. Percents of overlaps (**Overlap Percent**, OP) were defined as the proportion of a bin overlapped by an annotation region. **Distance** was defined as log2 transformed number of base pairs between between the centers of bins and regions defining genomic annotations. A log2-transformation of distance was used to account for the skewness present in the distributions of genomic distances (Supplementary Figure 1). These feature engineering procedures were performed to improve TAD boundary predictions by accounting for direct (OC), proportional (Signal/OP), and spatial (distance) relationships between TAD boundaries and genomic annotations.

@@

Expectedly, when using data as-is, that is, with class imbalance present, the models exhibited low MCCs. Similarly poor performances were found when using ROS as well. However, RUS and SMOTE re-sampling led to drastic improvement in performance, especially at higher resolutions. We found that RUS narrowly out performed SMOTE as the most optimal class balancing technique for all resolutions and predictor types (Figure 3). These results suggest that simple randomly under-sampling alone can drastically improve TAD boundary prediction. <!--These results suggest that a combination of both random over- and under-sampling (SMOTE) can improve TAD boundary prediction.-->Additionally, we found that at higher resolutions, random forest classifiers built using a distance-type feature space yielded substantially higher MCCs when compared to the peak signal stength, overlap count, and overlap percent predictor types (Figure 3). As with comparing class balancing techniques, this improvement was less evident at lower resolutions. <!--This indicated that models are better able to discriminate between TAD and non-TAD boundary regions when provided with spatially related associations.-->

@@

![**Models built using distance-type predictors combined with random under-sampling yielded optimal predictive performance.** Line plots comparing the performances of random forest classifiers built using different combinations of data level characteristics. Performances are compared across resolution, within each predictor-type: (A) OC, (B) OP, and (C) Distance, and between each re-sampling technique: no re-sampling (None; red), random over-sampling (ROS; green), random under-sampling (RUS; blue), and synthetic minority over-sampling (SMOTE; purple). Performances were evaluated using MCCs and aggregated by averaging across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/performances.png)

@@

We then aimed to compare the performances of genomic class-specifc models. In doing so, we inspected the proportion of predictive importance that was explained by each of the three classes of genomic elements including histone modifications (HM), cis-regulatory elements (CRE), and transcription factor binding sites (TFBS). For an optimally built random forest classifier at each data resolution--smote re-sampling, and distance-type predictors--we found that the majority of the proportional predictive importance was obtained by transcription factor binding sites for both cell lines (Supplementary Figure 3 A-B). Furthermore, models built solely on TFBS outperformed other class-specific models (Figure 4A; Supplementary Figure 4A). Notably, the TFBS-specific models outperformed models built on the combination of all genomic elements at lower resolutions, indicating that the inclusion of other types of annotations contributed to additional noise rather than signal in TAD boundary prediction. In summary, these results suggest that TFBS are the primary drivers of TAD boundary formation in both the GM12878 and K562 cell lines.

@@

Given that many TFBS are known to colocalize together on the linear genome and are therefore highly correlated, we sought to avoid overfitting by performing recursive feature elimination. We found that our models were able to obtain near-optimal performance using approximately eight TFBSs (Figure 4B; Supplementary Figure 4B). However, given that we train our models on chromosome-specific data, the most significant annotations vary for each chromosome. To determine the most common TFBSs, we clustered the predictive importance values across chromosomes (Figure 4C; Supplementary Figure 4C). We found that, similar to other studies [@sanborn2015chromatin; @fudenberg2016formation; @alipour2012self; @weintraub2017yy1; @gonzalez2017znf143], CTCF, RAD21, SMC3, and ZNF143 transcription factors were most predictive of TAD boundary regions.

![**Assessing top ranked transcription factors in TAD boundary prediction.** (A) Barplots comparing performances of models using specific classes of functional genomic elements as the feature space including histone modifications (HM; red), cis-regulated elements (CRE; green), transcription factor binding sites (TFBS; blue), in addition to a model with a feature space containing all three classes (Combined; purple). (B) Recursive feature elimination analysis used to reduce the number of TFBS to include in the optimally build random forest classifier. Error bars represent standard deviation from the mean MCC across each chromosome. (C) Clustered heatmap of predictive importances for the union of the top 8 most predictive TFBS across chromosomes.](Figures/annots_specifc_rfe_importances.png)

## Predictive modeling defines precise and biologically relevant location of TAD boundaries

After obtaining our optimally built random forest classifier, we investigated whether our predictive models can alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1 (Figure 5A). [??? Where is this figure? It seemed important; !!!added] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 5). [??? Need further drill down, as we discussed; !!!added distance to annotations figure in next paragraph]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!fixed] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 5B; Supplementary Figure 6).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs. Furthermore, when compared to called TAD boundaries, the boundaries that we predicted with our model were found to be statistically significantly closer to the top ranked TFBS, suggesting stronger enrichment profiles near predicted TAD boundaries (Supplementary Figure 7).

>>>>

Biological replicate Hi-C data were downloaded from GEO GSE63525 [@rao20143d] (Supplementary Table 1) for the GM12878 and K562 cell lines. The autosomal genomic coordinates (GRCh37/hg19 human genome assembly) of TAD boundaries were obtained from Hi-C data at 10 kb, 25 kb, 50 kb, and 100 kb resolution using the ARROWHEAD algorithm [@durand2016juicebox]. TAD boundaries called on chromosome 9 were omitted from all downstream analyses, for all resolutions and cell lines, due to the sparsity of the contact matrix at the 10 kb resolution for the K562 cell line. Genomic annotation data in the form of cell type-specific cis-regulatory elements (CRE), histone modifications (HM), and transcription factor binding sites (TFBS) were obtained from the UCSC Genome Browser Database (GRCh37/hg19 human genome assembly) [@Tyner:2017aa] (Supplementary Table 2).

## Computational framework for TAD boundary prediction using random forests

@@

where TP refers to the number of bins correctly identified as containing a TAD boundary (true positives), FP refers to the number of bins incorrectly identified as containing a TAD boundary (false positives), TN refers to the number of bins correctly identified as not containing a TAD boundary (true negatives), and FN refers to the number of bins incorrectly identified as not containing a TAD boundary (false negatives). Each of these quantities are obtained from the confusion matrix created by validating the model on the testing set. [???include reason for using mcc, why not others?]. We refer to the complete computational framework used to compare random forest model inputs as *TADRF*. <!--Additional performance metrics including AUROC, accuracy, sensitivity, and specificity were reported in the supplementary materials.-->

@@

Many genomic annotations, notably architectural proteins, tend to exhibit an extensive pattern of colocalization (correlation) [@van2014insulator]. To suitably reduce the feature space and prevent overfitting, we implemented recursive feature elimination to estimate the optimal number of necessary features to include in the best fitting *TADRF* model. The best performing subset was identified, ranging from 2 to the maximum number of features by powers of 2. We then aggregated the predictive importance of the union of the optimal set of features across chromosomes using the mean decrease in node impurity to determine the most common and top ranked annotations for predicting TAD boundaries. [??? TBD, at which stage it was done]

@@

Next, we aimed to investigate whether we could alleviate resolution limitations that plague conventional TAD calling algorithms. To do so, we extended our optimally built TAD prediction models to obtain predictions at base pair (bp) resolution. Naturally, when dealing with such fine resolution, there will exist collections of spatially close base pairs with similar predictor profiles (i.e. similar distances), and will thus yield similar predictions. To guard against inflating the number of predicted TAD boundaries we developed a technique that combines both hierarchical clustering and partitioning techniques to identify more precise TAD boundaries throughout the linear genome. We refer to our method as *preciseTAD*.

@@

We developed a computational framework, refered to as *TADRF*, aimed at determining the optimal set of data level characteristics necessary for TAD boundary prediction using Random Forest classification models. Our framework is resolution-specific and uses called TAD boundaries as ground truth to establish TAD-boundary and non-TAD boundary regions. Under our framework, we binned the linear genome at intervals specific to the Hi-C data resolution used to call TADs. TAD boundary regions were defined as regions on the linear genome where a genomic bin contained a called TAD boundary ($Y = 1$), while non-TAD boundary regions were defined as genomic bins that did not contain a TAD boundary ($Y = 0$) (Figure 1 A). Under this construction, the number of non-TAD boundary regions are exceed the number of TAD boundary regions. To cope with the imbalance among the class of genomic bins, we evaluated three re-sampling techniques (in addition to no re-sampling) before predicting TAD boundaries. *Random over-sampling* (ROS) was defined as sampling with replacement from TAD boundary regions. *Random under-sampling* (RUS) was defined as sampling with replacement from non-TAD boundary regions. Lastly, we considered *Synthetic minority over-sampling technique* (SMOTE), which is a combination of both random over- and under-sampling [@chawla2002smote].

Various one-dimensional cell line-specific ChIP-seq tracks (genomic elements) were used to construct the feature space for the random forests including cis-regulatory elements, histone modifications, and transcription factor binding sites. Four different feature engineering procedures were developed for each genomic element (Figure 1 B). *Peak Signal* was defined as the [???include definition]. Counts of overlaps (*Overlap Counts*, OC) were defined as the total number of genomic annotation regions that overlapped within a given genomic bin. Percents of overlaps (*Overlap Percent*, OP) were defined as the proportion of a bin overlapped by an annotation region. *Distance* was defined as log2 transformed number of base pairs between between the centers of bins and regions defining genomic annotations. A log2-transformation of distance was used to account for the skewness present in the distributions of genomic distances (Supplementary Figure 1). These feature engineering procedures were performed to improve TAD boundary predictions by accounting for direct (OC), proportional (Signal/OP), and spatial (distance) relationships between TAD boundaries and genomic annotations.

@@

Expectedly, when using data with class imbalance present, that is no re-sampling, the models exhibited low MCCs, with very little variability among different resolutions (Figure 3). Similarly poor performances were found when using ROS as well. However, RUS and SMOTE re-sampling led to drastic improvement in performance, especially at higher resolutions. We found that RUS narrowly out performed SMOTE as the most optimal class balancing technique for all resolutions and predictor types. These results suggest that simple randomly under-sampling alone can drastically improve TAD boundary prediction. <!--These results suggest that a combination of both random over- and under-sampling (SMOTE) can improve TAD boundary prediction.-->Additionally, we found that at higher resolutions, random forest classifiers built using a distance-type feature space yielded substantially higher MCCs when compared to the peak signal stength, overlap count, and overlap percent predictor types. As with comparing class balancing techniques, this improvement was less evident at lower resolutions. <!--This indicated that models are better able to discriminate between TAD and non-TAD boundary regions when provided with spatially related associations.-->

@@

![**Models built using distance-type predictors combined with random under-sampling at 10 kb yielded optimal predictive performance for GM12878.** Line plots comparing the performances of random forest classifiers built using different combinations of data level characteristics. Performances are compared across resolution, within each predictor-type: (A) OC, (B) OP, and (C) Distance, and between each re-sampling technique: no re-sampling (None; red), random over-sampling (ROS; green), random under-sampling (RUS; blue), and synthetic minority over-sampling (SMOTE; purple). Performances were evaluated using MCCs and aggregated by averaging across each chromosome, with error bars representing 1 standard deviation away from the mean.](Figures/performances.png)

@@

We then aimed to compare the performances of genomic class-specifc models. In doing so, we inspected the proportion of predictive importance that was explained by each of the three classes of genomic elements including histone modifications (HM), cis-regulatory elements (CRE), and transcription factor binding sites (TFBS). For an optimally built random forest classifier at each data resolution--smote re-sampling, and distance-type predictors--we found that the majority of the proportional predictive importance was obtained by transcription factor binding sites for both cell lines (Supplementary Figure 3 A-B). Furthermore, models built solely on TFBS outperformed other class-specific models (Figure 4A; Supplementary Figure 4A). Notably, the TFBS-specific models outperformed models built on the combination of all genomic elements at lower resolutions, indicating that the inclusion of other types of annotations contributed to additional noise rather than signal in TAD boundary prediction. Our results suggest that TFBS are the primary drivers of TAD boundary formation in both the GM12878 and K562 cell lines.

@@

Given that many TFBS are known to colocalize together on the linear genome and are therefore highly correlated, we sought to further optimize our TAD prediction model to avoid overfitting by performing recursive feature elimination. We found that, under our framework, we were able to obtain near-optimal performance using approximately eight TFBS (Figure 4B; Supplementary Figure 4B). However, given that we trained our model on chromosome-specific data, the most significant annotations were found to vary for each chromosome. To determine the most common TFBS, we clustered the predictive importances across chromosomes. We found that, similar to other studies [@sanborn2015chromatin; @weintraub2017yy1; @gonzalez2017znf143], CTCF, RAD21, SMC3, and ZNF143 transcription factors were most predictive of TAD boundary regions for all chromosomes (Figure 4C; Supplementary Figure 4C). Therefore, using our framework *TADRF* was able to identify the top molecular drivers of the 3D architecture of the human genome, components of the popular loop extrusion model [@fudenberg2016formation; @alipour2012self].

![**Assessing top ranked transcription factors in TAD boundary prediction for GM12878.** (A) Barplots comparing performances of models using specific classes of functional genomic elements as the feature space including histone modifications (HM; red), cis-regulated elements (CRE; green), transcription factor binding sites (TFBS; blue), in addition to a model with a feature space containing all three classes (Combined; purple). (B) Recursive feature elimination (RFE) analysis used to reduce the number of TFBS to include in the optimally built random forest model. Error bars represent standard deviation from the mean MCC across each chromosome. (C) Clustered heatmap of predictive importances for the union of the top 8 most predictive TFBS across chromosomes as determined from performing RFE. Importances were calculated as the mean decrease in accuracy among permuted features in out-of-bag samples.](Figures/annots_specifc_rfe_importances.png)

## *preciseTAD* defines precise and biologically relevant locations of TAD boundaries

After obtaining our optimally built *TADRF* model, we investigated whether we could leverage our model to alleviate resolution limitations of Hi-C data by predicting TAD boundaries at base pair resolution. Our method, referred to as *preciseTAD*, uses a combination of complete linkage hierarchical clustering with a scalable version of partitioning around mediods (PAM), known as clustering large applications (CLARA) to reduce many adjacent base pairs with similarly high predictive probabilities down to a set of the most representative and biologically meaningful predicted TAD boundaries (see Methods). *preciseTAD* was applied to chromosome-specific sections of the linear genome. For each chromosome, we defined our new validation data using distance-type predictors among only the transcription factor binding sites for each base pair. That is, each base pair had a corresponding row vector, **X**=$\{X_{1}, ..., X_{p}\}$, representing the $log_{2}$ genomic distances to the nearest respective transcription factor binding sites (Supplementary Table 2).

We first used ROC analysis to obtain an optimal probability threshold for defining potential predicted TAD boundaries. We reported an average threshold of ??? (SD=???) among the 21 autosomal chromosomes that were analyzed (mean=???, SD=??? for K562). We referred to collections of chromosome-specific base pairs that exceeded this threshold when used to validate our *TADRF* model as *predicted TAD boundary areas* (PTBA), given that collections of adjacent base pairs tended to exhibit similar prediction profiles. We found that there were a total of ??? PTBA, containing an average of ??? base pairs each (??? PTBA for K562; mean base pairs=???, SD=???). *preciseTAD* then implements complete linkage hierarchical clustering on the dissimilarity matrix constructed from the pairwise distances (in base pairs) among the midpoints of each chromsome-specific PTBA. Using the silhouette score (see methods), we reported a total of k=??? chromosome-specific clusters (k=??? for K562). We referred to these *k* clusters as *predicted TAD boundary regions* (PTBR). That is, each PTBR contained the midpoints of multiple PTBA. Lastly, given each PTBR, *preciseTAD* uses CLARA partitioning to obtain the most representative midpoint. We referred to this as the *predicted TAD boundary point* (PTBP). These PTBP were identified as the most representative base pair among each PTBR. *preciseTAD* was able to identify a total of ??? PTBP across each chromosome. We reported less predicted TAD boundaries than ARROWHEAD at 10 kb (Table 1). This is attributed to ARROWHEAD's inherent inflation of called TAD boundaries, especially near fuzzy areas along the diagonal of Hi-C contact matrices (Figure 5A). Likewise, under the framework of *preciseTAD*, any two TADs are inherently demarcated by only one predicted TAD boundary, leading to comparatively less TAD boundaries. We also find that predicted TAD boundaries identified by *preciseTAD* were much more likely to demarcate high concentrations of inter-chromosomal contacts among Hi-C contact matrices (Figure 5A). [???include a result that demonstrates this; discuss in meeting]

???stopped here; waiting to predicted tad boundaries results

<!--We were able to identify regions where the model predicted the base pair to be a TAD boundary with probability equal to 1 (Figure 5A). [??? Where is this figure? It seemed important; !!!added] By clustering the regions (see Methods), we were able to predict a comparable number of TAD boundary regions as called by the ARROWHEAD TAD caller. However, only 24% and 29% of predicted and called (and 10 kb flanked) TAD boundaries overlapped for the GM12878 and K562 cell line, respectively, suggesting differing locations in TAD boundaries throughout the linear genome (Supplementary Figure 5). [??? Need further drill down, as we discussed; !!!added distance to annotations figure in next paragraph]

For each predicted TAD boundary region, we were able to identify the summit locations of TAD boundary location by performing partitioning around medoids (see Methods). To compare the summit locations of predicted and called TAD boundaries, we evaluated the signal strength of ChIP-seq peak regions around a 10 kb flanking region. We found that the signal strength for the CTCF, RAD21, SMC3, and ZNF143 [??? Make all in this order!; !!!fixed] TFBS peaked at our predicted TAD boudaries, while was poorly defined around the called TAD boundaries (Figure 5B; Supplementary Figure 6).  Thus, our model is able to alleviate resolution limitations and precisely locate TAD boundaries that correspond to the highest signal strength of known TAD boundary-forming TFBSs. Furthermore, when compared to called TAD boundaries, the boundaries that we predicted with our model were found to be statistically significantly closer to the top ranked TFBS, suggesting stronger enrichment profiles near predicted TAD boundaries (Supplementary Figure 7).
-->
